Toxic Code Snippets on Stack Overflow

### 2018-05-14 cloverflow-tse/cloverflow.tex
Added QS clone ages and improved several sections in the paper.

\textbf{The visitors' survey}
The survey consists of 16 questions: 9 Likert's
scale questions, 3 yes/no questions, 2 multiple-choice questions, and 2
open-ended questions. 
The full survey can be found in our research note~\cite{Ragkhitwetsagul_RN2017}.
The first four questions are mandatory while the other 12
questions will be shown to the participants based on their previous answers. 
The survey collects information about the participant's software development
experience, the importance of Stack Overflow, reasons for reusing Stack Overflow
snippets, problems from Stack Overflow snippets, licensing of code on Stack
Overflow, and additional feedback. 
We adopted non-probability convenient sampling to invite participants for this
survey. Participation in the survey requires experience of visiting Stack
Overflow for solving programming tasks at least once. The participants were
invited to take the survey via five channels: social media post (Facebook),
\textsf{blognone.com}, a popular technology news and media community in Thailand, the University of Molise in Italy where the third author
works, \texttt{comp.lang.java.programmer} group and the Software Engineering
Facebook page. We received answers from 89 participants. Two participants never
copy code from Stack Overflow so we analysed the answers of the remaining 87
participants.
The survey was open for participation for 2 months, 
from 25 July 2017 to 25 September 2017, before we collected and analysed
the responses.

@@

\vspace{0.25cm}
\textit{RQ 1.1 How often are Stack Overflow
	answerers aware of the outdated code and 
	licensing conflicts when
	they answer a question on Stack Overflow?} 
\vspace{0.25cm}

@@

\begin{boxquote}
	To answer RQ 5.1, regarding outdated code, although most of the Stack Overflow answerers are aware that their code
	can be outdated, 51.5\% of the answerers were very rarely notified and 35.5\% have never been notified of
	outdated code in the answers. After being notified, 19.8\% of them rarely or never fix the outdated code.

@@

@@

@@

@@

%We discovered two issues regarding code snippets on Stack Overflow in this
% study: (1)~outdated online clones on Stack Overflow and (2)~missing or
% incompatible license in Stack Overflow snippets.
%that there is a considerable amount of cloned code between Stack Overflow that
% are cloned from 111 projects in the Qualitas corpus. 64 clone pairs are
% outdated and 358 snippets have incompatible software licenses to their
% originals.
Our study discovers links from code in open source projects to code snippets on
Stack Overflow using clone detection techniques. These links enable us to
discover toxic code snippets with outdated code or licensing problems. 
The links can be exploited further to mitigate the problems of reusing outdated online clones and
incompatible license on Stack Overflow code snippets. We propose the following
actionable items:

\begin{itemize} 
	\item \textbf{Preventive measure:} We
	encourage Stack Overflow to enforce attribution when source code snippets have
	been copied ``from'' licensed software projects to Stack Overflow. Moreover, an
	IDE plug-in that can automatically detect pasted source code and follow the link
	to Stack Overflow and then to the original open source projects, could also
	prevent the issue of license violation. 
	\item \textbf{Detective measure:} A
	system to detect outdated source code snippets on Stack Overflow is needed. The
	system can leverage the online clone detection techniques in this study to
	periodically check if the cloned snippets are still up-to-date with their
	originals. %Moreover, Stack Overflow posters have to provide the information of
	% the project including the original file name and its online repository (if
	% exists).
	With such a system, the poster can be notified when the code has been updated
	in the original project so that he/she can update their code on Stack Overflow
	accordingly. On the other hand, with a crowdsourcing solution using an IDE
	plug-in, developers can also report the corrected version of outdated code back
	to the original Stack Overflow threads when they reuse outdated code and make
	corrections to them. 
\end{itemize}

>>>>

\textbf{The visitors' survey}
The survey consists of 16 questions: 9 Likert's
scale questions, 3 yes/no questions, 2 multiple-choice questions, and 2
open-ended questions. 
The full survey can be found in our research note~\cite{Ragkhitwetsagul_RN2017}.
The first four questions are mandatory while the other 12
questions will be shown to the participants based on their previous answers. 
The survey collects information about the participant's software development
experience, the importance of Stack Overflow, reasons for reusing Stack Overflow
snippets, problems from Stack Overflow snippets, licensing of code on Stack
Overflow, and additional feedback. 
We adopted non-probability convenient sampling to invite participants for this
survey. Participation in the survey requires experience of visiting Stack
Overflow for solving programming tasks at least once. The participants were
invited to take the survey via five channels: social media post (Facebook),
\textsf{blognone.com}, a popular technology news and media community in Thailand, the University of Molise in Italy where the third author
works, \texttt{comp.lang.java.programmer} group and the Software Engineering
Facebook page. 
The survey was open for participation for 2 months, 
from 25 July 2017 to 25 September 2017, before we collected and analysed
the responses.

@@

\vspace{0.25cm}
\textit{How often are Stack Overflow
	answerers aware of the outdated code and 
	licensing conflicts when
	they answer a question on Stack Overflow?} 
\vspace{0.25cm}

@@

\begin{boxquote}
	To answer RQ 1.1, regarding outdated code, although most of the Stack Overflow answerers are aware that their code
	can be outdated, 51.5\% of the answerers were very rarely notified and 35.5\% have never been notified of
	outdated code in the answers. After being notified, 19.8\% of them rarely or never fix the outdated code.

@@

\subsubsection{The Visitor Survey}

We received answers from 89 participants. 
Two participants never
copy code from Stack Overflow so we analysed the answers of the remaining 87
participants.

\textbf{General Information: }
Twenty four (27\%) and twenty one (24\%) participants have over 10 years and 5--10
years of experience respectively. There are 19 participants (21\%) who have 3--5
years, 18 (20\%) who have 1-2 years, and 7 (8\%) participants who have less than a
year of programming experience.

\textbf{Choice for programming solutions: }
Stack Overflow is ranked higher than official documentation, online
repositories, and books as the resource to look for programming solutions.
Developers rely on Stack Overflow answers because they are easy to search for on
the web. Moreover, 64\% of the participants reuse code snippets from Stack
Overflow at least once a week. They copy code from Stack Overflow because they
can be found easily from search engine, solve similar problems to their
problems, provide helpful context, and offer voting and accepted answers.

\vspace{0.25cm}
\textit{Do Stack Overflow
	visitors experience outdated code and check for licensing conflicts when
	they copy and reuse answer from Stack Overflow?}
\vspace{0.25cm}

\begin{table}
	\centering
	\caption{Problems from Stack Overflow code snippets}
	\label{tab:visitor_survey_code_problems}
	\begin{tabular}{lr}
		\toprule
		Problem & Amount \\
		\midrule
		Mismatched solutions & 40 \\
		Outdated solutions & 39 \\
		Incorrect solutions	& 28 \\
		Buggy code & 1 \\
		\bottomrule
	\end{tabular}
\end{table}

\begin{table}
	\centering
	\caption{Frequency of reporting the problems to Stack Overflow posts}
	\label{tab:visitor_survey_report_problem}
	\begin{tabular}{lrr}
		\toprule
		Report? & Amount & Percent \\
		\midrule
		Very Frequently (81--100\% of the time)	& 1 & 1.8\% \\
		Frequently (61--80\% of the time) & 1 & 1.8\% \\
		Occasionally (41--60\% of problematic snippets) & 3 & 5.3\% \\
		Rarely (21--40\% of problematic snippets) & 8 & 14.0\% \\
		Very rarely (1--20\% of problematic snippets) & 8 & 14.0\% \\
		Never (0\% of problematic snippets) & 36 & 63.2\% \\
		\midrule
		Total & 57 & 100\% \\
		\bottomrule
	\end{tabular}
\end{table}

The survey results show that 57 out of 87 Stack Overflow visitors encountered a
problem from reusing Stack Overflow code snippets. Ten participants experienced
problems for more than 80\% of the copied snippets and sixteen participants
faced problems for 40--60\% of the reused code. As shown in~\Cref{tab:visitor_survey_code_problems}, 
the problems ranked by frequency
include mismatched solutions (40), outdated solutions (39), incorrect solutions
(28), and buggy code (1). Sixty-three percent of the participants never report
the problems back to Stack Overflow (\Cref{tab:visitor_survey_report_problem}). 
The ways of reporting the problems
(22 answers) included down-voting the answer containing the problematic code
snippet (8), writing a comment saying that the code has problems (10),
contacting the answerers regarding the problems directly (2), and posting a
better snippet as new answer on same topic (2).

\begin{table}
	\centering
	\caption{Check for licensing conflicts before using Stack Overflow snippets}
	\label{tab:visitor_survey_license_check}
	\begin{tabular}{lrr}
		\toprule
		License check? & Amount & Percent \\
		\midrule
		Very frequently (81--100\% of the time) & 0 & 0.0\% \\
		Frequently (61--80\% of the time) & 7 & 8.1\%
\\
		Occasionally (41--60\% of the time) & 6 & 6.9\%
\\
		Rarely (21--40\% of the time) & 6 & 6.9\% \\
		Very rarely (1--20\% of the time) & 11 & 12.6\% \\
		Never (0\% of the time)	& 57 & 65.5\% \\
		\midrule
		Total & 87 & 100\% \\
\bottomrule
\end{tabular}
\end{table}

In addition, 74 out of the 87 (85\%) participants are not
aware of Stack Overflow CC BY-SA 3.0 license and 62\% never give attributions to
the Stack Overflow posts they copied the code snippets from. 
As shown in \Cref{tab:visitor_survey_license_check}, we found that 66\%
of the visitors never check for software licensing conflicts between Stack
Overflow code snippets and their projects. Interestingly, 9\% of the
participants encountered legal issues. Unfortunately, due to the anonymity of
the survey, we could not investigate further regarding the legal issues that the
participants faced from using Stack Overflow code snippets.

The visitors survey confirms the findings from the previous studies that Stack
Overflow code snippets can be problematic~\cite{Zhang2018,Acar2016,An2017,Fischer2017}. Sixty-six
percent of the visitors experienced a problem from reusing Stack Overflow code
snippets ranging from incorrect solutions, outdated solutions, mismatched
solutions, to buggy code. Although they are aware of the problems, half of them
(56\%) never reports back to the Stack Overflow discussions. On the other hand,
the visitors rarely give attributions to Stack Overflow when they reuse code
snippets from the website, similar to the findings reported
by~\cite{Baltes2017}. The visitors are generally not aware of the CC BY-SA 3.0
license, and more than half of them never check for license compatibility when
reusing Stack Overflow code snippets. 
Sixty-nine Stack Overflow visitors (79\%) who adopted code from Stack Overflow never 
check if the code snippet originated from a different source (e.g. an open source project) 
with an incompatible license to their projects.
Moreover, we also found that 9\% of the participants
encountered legal issues by copying code from Stack Overflow. To the best of our
knowledge, a study of legal problems from reusing Stack Overflow code snippets
has never been done before and will be our future work.

\begin{boxquote}
	To answer RQ 1.2, Stack Overflow visitors experienced several issues from Stack
Overflow answers including outdated code. The majority of them are not aware of
CC BY-SA 3.0 license enforced by Stack Overflow and never check for license
conflicts when reusing code snippets.
\end{boxquote}

@@

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{boxplot_clone_age}
	\caption{Age of online code clones.}
	\label{fig:boxplotcloneage}
\end{figure}

@@

To check occurrences of our Stack Overflow outdated code snippets in software projects, 
we

@@

%We discovered two issues regarding code snippets on Stack Overflow in this
% study: (1)~outdated online clones on Stack Overflow and (2)~missing or
% incompatible license in Stack Overflow snippets.
%that there is a considerable amount of cloned code between Stack Overflow that
% are cloned from 111 projects in the Qualitas corpus. 64 clone pairs are
% outdated and 358 snippets have incompatible software licenses to their
% originals.

This study empirically show from the surveys and the clone detection experiment that
online code clones occur on Stack Overflow and the clones may become toxic due
to outdated code and software license incompatibility. The findings lead to the
insight about toxicity of online code clones and two actionable items
we are proposing to the Stack Overflow community.

\subsubsection{Toxicity of outdated and license-incompatible clones}
We found only a small amount of toxic outdated code snippets in open source projects on
GitHub. Besides 12 buggy and outdated code snippets found in 12 projects, the
rest were non-harmful clones of the outdated code. Although other studies show
that Stack Overflow code snippets may contain security
vulnerabilities~\cite{Acar2016,Fischer2017} or API misuse~\cite{Zhang2018}, we
found in this study that the damage caused by reusing outdated code on Stack
Overflow is not high.

On the other hand, the missing of licensing statement of online code clones on
Stack Overflow can cause more damage than outdated code. As shown in our study and
also in the study by An et al. \cite{An2017}, some online clones on Stack Overflow are
originally licensed under more restrictive license than Stack Overflow's CC
BY-SA 3.0. If these missing-license online clones are reused in a software with
incompatible license, the software owner may face legal issues. The software
auditing service such as Black Duck Software or nexB, which can effectively
check for license compliance of code copied from open source projects, cannot
check for the original license of the cloned code snippets on Stack Overflow. Although
Stack Overflow answerers participated in our survey believe that most of the
code snippets on Stack Overflow are too small to claim for copyright and they
fall under fair-use, there is still a risk due to different legal system in each
country. For example, Germany legal system does not have a concept of fair use~\cite{?}
or the number of minimum lines of code to be consider trivial, i.e.~de minimis,
is also different from case to case or from country to country~\cite{?}.

\subsubsection{Actionable Items}
Our study discovers links from code in open source projects to code snippets on
Stack Overflow using clone detection techniques. These links enable us to
discover toxic code snippets with outdated code or licensing problems. 
The links can be exploited further to mitigate the problems of reusing outdated online clones and
incompatible license on Stack Overflow code snippets. We propose the following
actionable items:

\textbf{Preventive measure:} We encourage Stack Overflow to enforce attribution
when source code snippets have been copied ``from'' licensed software projects
to Stack Overflow. Moreover, an IDE plug-in that can automatically detect pasted
source code and follow the link to Stack Overflow and then to the original open
source projects, could also prevent the issue of license violation. We foresee
the implementation of the IDE plugin using a combination of scalable code clone
detection~\cite{Sajnani2016} or clone search techniques~\cite{Kim2018} and
automated software licensing detection~\cite{German2010}. In this study, we
performed the check using a set of code clone detectors (Simian and SourcererCC)
and software licensing detector (Ninka) but we had to operate the tools
manually. Using the knowledge obtained from this study, we plan to build an
automated and scalable clone search with licensing detection. With the proposed
solution, we can use the tool to create a database of code snippets on Stack
Overflow and allow the users to search for clones and check their licenses. The
clone search tool can offer a service via REST API and integrated into the IDE
plugin. Every time a new code fragment is pasted into the IDE, the plugin
performs the check by calling the clone search tool service and report the
finding to the developers in real time.

In addition, we also performed a study of two open source software auditing
platforms/services:~BlackDuck
Software\footnote{https://www.blackducksoftware.com} and
nexB\footnote{https://www.nexb.com}. For BlackDuck Software, we found from their
report~\cite{CORSI2017} that while they check for code copied from open source
projects including GitHub and Stack Overflow and analyse their license
compatibility with their customer software, the BlackDuck auditing system will
treat the code snippets on Stack Overflow as having ``unknown'' license because
it does not know the original license of Stack Overflow code snippets. For nexB,
their product does not mention checking of reused source code from Stack
Overflow. So, our proposed service, which can offer more precise licensing
information of Stack Overflow code snippets, will be useful as an add-on license
check for code copying from Stack Overflow.

\textbf{Detective measure:} A
system to detect outdated source code snippets on Stack Overflow may be needed. The
system can leverage the online clone detection techniques in this study to
periodically check if the cloned snippets are still up-to-date with their
originals.

While checking if the copied code snippets on Stack Overflow are still up-to-date with the latest version of their originals can possibly be done automatically, it is a challenging task to automate the process of establishing the origin of code clones. As shown in this study, we still relied on manual validation to read the post, understand the context, and use multiple hints based on human judgement (comments in the code, natural text in the question and answers, date/time of the posts) to conclude that the code snippets are actually copied from Qualitas to Stack Overflow. Thus, one viable solution for now is encouraging the Stack Overflow developers to always include the origin of the copied code snippet in the post so that this link is always established at the posting time. Even better, Stack Overflow can provide an optional form to fill in when an answerer post an answer if he or she copies code from other software projects. The form should include the origin of the code snippet (possibly as a GitHub url) and its original license. Using this manually established links at posting time, we can then automate the process of checking for outdated code.

With such a system, the poster can be notified when the code has been updated
in the original project so that he/she can update their code on Stack Overflow
accordingly. On the other hand, with a crowdsourcing solution using an IDE
plug-in, developers can also report the corrected version of outdated code back
to the original Stack Overflow threads when they reuse outdated code and make
corrections to them.
